---
title: "P8131-HW5-yz4184"
author: "Yunlin Zhou"
date: "3/24/2022"
output: pdf_document
---
```{r, echo = T, message = FALSE, results='hide', warning=FALSE}
library(tidyverse)
library(pscl)
```

### Import the data

```{r, message = FALSE, results='hide', warning=FALSE}
crab_df = read.csv("./HW5-crab.txt", sep = "")

para_df = read.csv("./HW5-parasite.txt", sep = "")%>%
  janitor::clean_names()%>%
  mutate(area = as.factor(area),
        year = as.factor(year)) %>%
  select(year, intensity, length, area)
```

# Question 1

## (a)

###  Fit a Poisson model (M1) with log link with W as the single predictor.

```{r}
M1 <- glm(Sa ~ W,
          family = poisson(link = log),
          data=crab_df)

summary(M1) 
```

The log rate ratio of satellites number would increase 0.16405 as carapace width increases 1 unit. The deviance of this model is 567.88 with 171 df. Carapace width is statistically significant.

### goodness of fit

```{r}
# pearson residual
P1 = sum(residuals(M1, type = 'pearson')^2)

p_val1 = 1 - pchisq(P1, df = 171)
p_val1
```

Since the p-value of the goodness fit is less than 0.5, we reject the null hypothesis that the M1 is a good fit.

## (b)

###  Fit a model (M2) with W and Wt as predictors.

```{r}
M2 <- glm(Sa ~ W + Wt,
          family = poisson(link = log),
          data=crab_df)

summary(M2)
```

The log rate ratio of satellites number would increase 0.04590 as carapace width increases 1 unit, holding other variables fixed.

The log rate ratio of satellites number would increase 0.44744 as weight increases 1 unit, holding other variables fixed.

The deviance of this model is 559.89 with 170 df. Weight is statistically significant.

### Compare M1 with M2

```{r}
# analysis of deviance
D1 = M1$deviance - M2$deviance

p_val2 = 1 - pchisq(D1, 1)
p_val2
```

Since the p-value of the goodness fit is less than 0.5, we reject the null hypothesis that the M1 is better.


## (c)

### Check over dispersion in M2.

```{r}
# Pearson residual
P2=sum(residuals(M2, type = 'pearson')^2)

phi=P2/170
phi

summary(M2, dispersion = phi)
```

Since phi is 3.156449 which is greater than 1, we can conclude that there is over-dispersion in this model.

After adjusting for over dispersion, the standard error of carapace width changes from 0.04677 to 0.08309; the standard error of weight changes from 0.15864 to 0.28184. 

The log rate ratio of satellites number would increase 0.04590 as carapace width increases 1 unit, holding other variables fixed.

The log rate ratio of satellites number would increase 0.44744 as weight increases 1 unit, holding other variables fixed.

The deviance of this model is 559.89 with 170 df.

\newpage

# Question 2

## (a)

###  Fit a Poisson model with log link to the data with area, year, and length as predictors.

```{r}
M3 = glm(intensity ~ area + length + year,
              family = poisson(link = log),
              data = para_df)

summary(M3)
```

The log rate ratio of intensity would decrease 0.2119557 as the area changes from 1 to 2, holding other variables fixed.

The log rate ratio of intensity would decrease 0.1168602 as the area changes from 1 to 3, holding other variables fixed.

The log rate ratio of intensity would increase 1.4049366 as the area changes from 1 to 4, holding other variables fixed.

The log rate ratio of intensity would decrease 0.0284228 as length increases 1 unit, holding other variables fixed.

The log rate ratio of intensity would increase 0.6702801 as the year changes from 1999 to 2000, holding other variables fixed.

The log rate ratio of intensity would decrease 0.2181393 as the year changes from 1999 to 2001, holding other variables fixed.

## (b)

### Test for goodness of fit of the model.

```{r}
P3 = sum(residuals(M3, type = 'pearson')^2)

p_val3 = 1 - pchisq(P3, df = 1184)
p_val3
```

Since the p-value of the goodness fit is less than 0.5, we reject the null hypothesis that the M3 is a good fit.

## (C)

### Fit zero-inflated poisson regression model

```{r}
M4 <- zeroinfl(intensity ~ area + length + year ,
               data = para_df)
summary(M4)
```

### Within the susceptible fish, we use the poisson model:

The log rate ratio of intensity would increase 0.2687838 as the area changes from 1 to 2, holding other variables fixed.

The log rate ratio of intensity would increase 0.1463174 as the area changes from 1 to 3, holding other variables fixed.

The log rate ratio of intensity would increase 0.9448070 as the area changes from 1 to 4, holding other variables fixed.

The log rate ratio of intensity would decrease 0.0368067 as length increases 1 unit, holding other variables fixed.

The log rate ratio of intensity would increase 0.3919828 as the year changes from 1999 to 2000, holding other variables fixed.

The log rate ratio of intensity would decrease 0.0448457  as the year changes from 1999 to 2001, holding other variables fixed.


### Since we donot know which fish are susceptible, we assume that all the variables are realted to binomial model.

The log rate ratio of intensity would increase 0.718680 as the area changes from 1 to 2, holding other variables fixed.

The log rate ratio of intensity would increase 0.657710 as the area changes from 1 to 3, holding other variables fixed.

The log rate ratio of intensity would decrease 1.022864 as the area changes from 1 to 4, holding other variables fixed.

The log rate ratio of intensity would decrease 0.009889 as length increases 1 unit, holding other variables fixed.

The log rate ratio of intensity would decrease 0.752121 as the year changes from 1999 to 2000, holding other variables fixed.

The log rate ratio of intensity would increase 0.456533  as the year changes from 1999 to 2001, holding other variables fixed.
