---
title: "P8131-HW5-yz4184"
author: "Yunlin Zhou"
date: "3/24/2022"
output: pdf_document
---
```{r, echo = T, message = FALSE, results='hide', warning=FALSE}
library(tidyverse)
```

### Import the data

```{r, message = FALSE, results='hide', warning=FALSE}
crab_df = read.csv("./HW5-crab.txt", sep = "")

para_df = read.csv("./HW5-parasite.txt", sep = "")%>%
  janitor::clean_names()%>%
  mutate(area = as.factor(area),
        year = as.factor(year)) %>%
  select(year, intensity, length, area)
```

# Question 1

## (a)

###  Fit a Poisson model (M1) with log link with W as the single predictor.

```{r}
M1 <- glm(Sa~W, family = poisson(link = log), data=crab_df)
summary(M1) 
```

The log rate ration of satellites number would change 0.16405 as carapace width changes 1 unit. The deviance of this model is 567.88 with 171 df. Carapace width is statistically significant.

### goodness of fit

```{r}
# pearson residual
P1 = sum(residuals(M1, type = 'pearson')^2)
p_val1 = 1 - pchisq(P1, df = 171)
p_val1
```

Since the p-value of the goodness fit is less than 0.5, we reject the null hypothesis that the M1 is a good fit.

## (b)

###  Fit a model (M2) with W and Wt as predictors.

```{r}
M2 <- glm(Sa~W+Wt, family = poisson(link = log), data=crab_df)
summary(M2) 
```

The log rate ration of satellites number would change 0.04590 as carapace width changes 1 unit; and would change 0.44744 as weight changes 1 unit. The deviance of this model is 559.89 with 170 df. Weight is statistically significant.

### Compare M1 with M2

```{r}
# analysis of deviance
D1 = M1$deviance - M2$deviance
p_val2 = 1 - pchisq(D1, 1)
p_val2
```

Since the p-value of the goodness fit is less than 0.5, we reject the null hypothesis that the M1 is better.


## (c)

### Check over dispersion in M2.

```{r}
P2=sum(residuals(M2, type = 'pearson')^2)
phi=P2/170
phi

summary(M2, dispersion = phi)
```

Since phi is 3.156449 which is greater than 1, we can conclude that there is over-dispersion in this model.

After adjusting for over dispersion, the standard error of carapace width changes from 0.04677 to 0.08309; the standard error of weight changes from 0.15864 to 0.28184. 

The log rate ration of satellites number would change 0.04590 as carapace width changes 1 unit; and would change 0.44744 as weight changes 1 unit. The deviance of this model is 559.89 with 170 df. Weight is statistically significant.

